Supplements for ``Tensor denoising and completion based on ordinal observations''

### 2020-02-11 ICML/arxiv/ordinalT_supp_v3.tex
upload

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}
\externaldocument{ordinalT}
\input macros.tex

@@

\section{Proofs}
\subsection{Estimation error for tensor denoising}
\begin{proof}[Proof of Theorem~\ref{thm:rate}]
We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$. It follows from the expression of $\flogl(\Theta)$ that
\begin{align}\label{eq:property}
{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
\end{align}
for all $\omega\in[d_1]\times \cdots \times [d_K]$.
Define $d_{\text{total}}=\prod_k d_k$. Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$. Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector. By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix. Recall that
\[
U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{\dot{g}_\ell(\theta)\over g_\ell(\theta)}>0 \quad \text{and}\quad
L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
\]
Therefore, all entries in $\fplogl$ are upper bounded $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.

@@

\subsection{Sample complexity for tensor completion}
\begin{proof}[Proof of Theorem~\ref{thm:minimax}]

@@

@@

@@

\section{Convexity of the likelihood function}
\begin{thm}\label{thm:convexity}
\begin{align}
 \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\Big\}, \text{ where } f(x) = \frac{e^x}{1+e^x}.
 \end{align}
is a concave function to $(\Theta,\mb)$
\end{thm}
\begin{proof}
We use $(\Theta,\mb)$ abusively as $(Vec(\Theta)^T,\mb^T)^T$ in this proof.
Let us denote $\lambda(u,v) = \log \big[f(u)-  f(v)\big]$.
It is enough to show that $\lambda(u,v)$
is a concave function to $(u,v)$ where $u>v$. Because if $\lambda(u,v)$ is a concave function and $u,v$ are both linear functions of $(\Theta,\mb) $ such that
$$u = \ma_1^T(\Theta,\mb), v = \ma_2^T(\Theta,\mb) \text{ for some }\ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1},$$ then $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ma_2^T(\Theta,\mb)))$ is a concave function to $(\Theta,\mb)$
by the definition of the convexity. Furthermore, we can conclude that $\logl(\Theta, \mb) $ is a concave function to $(\Theta,\mb)$ because $\logl(\Theta, \mb)$
can be written as the form of summations of $\lambda(u,v)$. To prove the convexity of $\lambda(u,v)$, we have the following.
\begin{equation*}
\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{(u,v)}(x)f'(x)dx\big].
\end{equation*}
Notice that  $\mathds{1}_{(u,v)}(x)$ is log-concave to $(u,v,x)$ and $f'(x)$ is log-concave to $x$ because $f'(x) = \frac{e^x}{(1+e^x)^2}$. Those convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
is a log-concave function to $(u,v,x)$. Therefore, we can conclude that the $\lambda(u,v)$ is a concave function from the lemma~\ref{lem:lossconvexity}.
\end{proof}

\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let $$G(x) = \int_{\mathbb{R}^n}F(x,y)dy.$$ If $F(x,y)$ is a log concave function to (x,y), then $G(x)$ is a log concave function.
\end{lem}

\section{Brain clustering method and result}
\subsection{Clustering method}
We perform clustering based on the estimated latent parameter $\hat\Theta$. Our clustering method is based on multidimensional version of Principal Component Analysis (PCA) and Singular Value Decomposition (SVD). In matrices case, we perform clustering on a $m\times n$ data matrix $X$ based on the following procedure. First, we factorize $X$ into
\begin{equation*}
X = U\Sigma V^T,
\end{equation*}
where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns. We interpret each column of $V$ as a principal direction or axis and $U\Sigma$ as principal components. From this interpretation, we perform $K$-means method to $m$ rows of $U\Sigma$. We suggest to apply this procedure to the estimated latent parameter tensor $\hat\Theta$. For the notational convenience, we use $\Theta$ as $\hat\Theta$ in this section. From the Tucker decomposition which is multidimensional version of SVD, we have
\begin{equation*}
\Theta = \tC\times_1\mM_1\times_2\cdots\times_K\mM_K,
\end{equation*}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}. We perform mode-$k$ matricization of the tensor $\Theta$ to cluster data from mode-$k$ variable as follows.
\begin{equation*}
\Theta_{(k)} = \mM_k\tC_{(k)}\big(\mM_K\otimes\cdots\otimes\mM_1\big)
\end{equation*}
We interpret $\big(\mM_K\otimes\cdots\otimes\mM_1\big)$ as principal axes and $\mM_k\tC_{(k)}$ as principal components. By putting other mode variables as principal axes, we can exclude effects from those variables and can see $k$-th mode effect from checking principal components, $\mM_k\tC_{(k)}$. Therefore, we perform $K$-means method to $d_k$ rows of the principal components like in the matrices case.

\subsection{Clustering result on HCP}
We perform $K$-means method to cluster brain nodes based on the suggested above method. We have the estimated latent parameter $\hat\Theta$ with the rank $(23,23,8)$ which minimizes the BIC value. The number of clusters is chosen to be eight from the elbow method in figure ~\ref{figure:elbow}.
The table in \ref{table:clustering} shows eight cluster groups and their entries. Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain. From the group $\#3$ to the group $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain. For example, the group $\#3$ and the group $\#4$ are the left and the right side of the regions called Supramarginal gyrus. Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}. The figure \ref{figure:brain image} shows the nodes in the dataset on the brain image. Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain. The results show that the clustering based on our model successfully group the brain nodes without knowledge of the brain region.
\begin{figure}[H]
  \label{figure:elbow}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
  \end{center}
  \caption{The elbow plot for the number of clusters}
\end{figure}

\begin{table}[H]
  \label{table:clustering}
  \center
  \begin{small}
    \footnotesize
  \begin{sc}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
\\ \hline
Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
\end{tabular}
\end{sc}
\end{small}
\caption{Result of the brain nodes clustering: the last alphabet in each node name means the left side for `L' and the right side for `R' in the brain}
\end{table}

\begin{figure}[H]
  \centering
 \includegraphics[width = 0.45\textwidth]{brain1.jpg}
 \includegraphics[width = 0.45\textwidth]{brain2.jpg}
 \includegraphics[width = 0.45\textwidth]{brain3.jpg}
 \includegraphics[width = 0.45\textwidth]{brain4.jpg}
  \label{figure:brain image}
  \caption{Brain image of eight clusters: entries of each cluster are connected by edges}
\end{figure}

>>>>

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}
\externaldocument{ordinalTv3}
\input macros.tex
\setlength\parindent{0pt}
\allowdisplaybreaks

@@

\section{Proofs}
\subsection{Estimation error for tensor denoising}
\begin{proof}[Proof of Theorem~\ref{thm:rate}]
We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$. It follows from the expression of $\flogl(\Theta)$ that
\begin{align}\label{eq:property}
{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
\end{align}
for all $\omega\in[d_1]\times \cdots \times [d_K]$.
Define $d_{\text{total}}=\prod_k d_k$. Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$. Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector. By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix. Recall that
\begin{equation}\label{eq:bound}
U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
\end{equation}
Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.

@@

\begin{proof}[Proof of Theorem~\ref{thm:minimax}]

@@

\subsection{Sample complexity for tensor completion}

@@

\subsection{Convexity of the log-likelihood function}
\begin{thm}\label{thm:convexity}
Define the function
\begin{align}\label{eq:function}
 \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\right\},
 \end{align}
where $f(\cdot)$ satisfies Assumption~\eqref{eq:ass}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$. 
\end{thm}

\begin{proof}
Define $d_{\text{total}}=\prod_k d_k$. With a little abuse of notation, we use $(\Theta,\mb)$ to denote a length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. Let us denote a bivariate function
\begin{align}
\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
\end{align}
It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.

Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
\[
u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
\]
Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity. Furthermore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.

Now, we prove the concavity of $\lambda(u,v)$. Note that
\begin{equation*}
\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
\end{equation*}
where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere. Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$ and $f'(x) = \frac{e^x}{(1+e^x)^2}$ is log-concave in $x$. The convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$. 
\end{proof}

\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let 
\[
G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
\]
If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
\end{lem}

@@

\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}

When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
\begin{equation}\label{eq:joint}
(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\logl(\Theta,\mb),
\end{equation}
where 
\[
\tP=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\}, \quad \tB=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
\]
The estimation accuracy is assessed using the mean squared error (MSE):
\[
\text{MSE}(\hat \Theta,\Theta)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\Theta},\quad \text{MSE}(\hat \mb,\mb)={1\over L-1}\FnormSize{}{\hat \mb-\mb}.
\]

We introduce several quantities that will used in our theory:
\begin{enumerate}
\item We make the convention that $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$, $f(\infty)=1$, and $\dot{f}(-\infty)=\ddot{f}(-\infty)=\dot{f}(\infty)=\ddot{f}(\infty)=0$.

\item The difference function $g_\ell(\theta)$ is defined as $g_\ell(\theta)=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\theta\in\mathbb{R}$ and $\ell\in[L]$.

\item Define $n_\ell=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell\}}$, i.e., the number of tensor entries taking value on $\ell\in[L]$.

\item With a little abuse of notation, we re-define the constants in~\eqref{eq:bound} as
\begin{equation}\label{eq:UL}
U_{\alpha,\beta,\Delta}=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}, \quad \text{and}\quad
L_{\alpha,\beta,\Delta}=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
\end{equation}
\item We define two additional constants:
\begin{align}\label{eq:CD}
C_{\alpha,\beta,\Delta} &= \max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}  \max_{\ell\in[L-1]}\max\left\{\frac{\dot{f}(\theta+b_\ell)}{g_\ell(\theta)},\ \frac{\dot{f}(\theta+b_\ell)}{g_{\ell+1}(\theta)}\right\},\notag \\
D_{\alpha,\beta,\Delta} &=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \min_{\ell\in[L-1]}\min\left\{ - {\partial \over \partial \theta}\left({\dot{f}(\theta+b_{\ell})\over g_\ell(\theta)}\right),\ {\partial \over \partial \theta}\left({\dot{f}(\theta+b_{\ell})\over g_{\ell+1}(\theta)}\right)\right\}\notag \\
&=  \min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]}\min \left\{-\frac{\ddot{f}(\theta+b_\ell)g_\ell(\theta)-\dot{f}(\theta+b_\ell)\dot{g}_\ell(\theta)}{g^2_\ell(\theta)},\ \right. \notag \\
&\hspace{1.5in} \left. \frac{\ddot{f}(\theta+b_\ell)g_{\ell+1}(\theta)-\dot{f}(\theta+b_\ell)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\}.
\end{align} 
\end{enumerate}

We make the following assumptions about the link function.  
\begin{assumption}\label{ass:joint}
The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
\begin{enumerate}
\item $f(\theta)$ is twice-differentiable and strictly increasing in $\theta$.
\item $\dot{f}(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$. 
\item The function ${\dot{f}(\theta+b_\ell)\over g_\ell(\theta)}$ is strictly decreasing with respect to $\theta$ for all $\mb\in\tB$. 
\item The function ${\dot{f}(\theta+b_\ell)\over g_{\ell+1}(\theta)}$ is strictly increasing with respect to $\theta$ for all $\mb\in\tB$. 
\end{enumerate}
\end{assumption}

\begin{rmk}
The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(\theta+b_{\ell})-f(\theta+b_{\ell-1})$. Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are well-defined. Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$. Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$.
\end{rmk}
\begin{rmk}
In particular, for logistic link $f(x) = \frac{1}{1+e^{-x}}$ , we have 
\begin{align}
C_{\alpha,\beta,\Delta}&=\max_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\max_{\ell\in[L-1]}\max
\left\{
{ 1\over 1-e^{-(b_{\ell}-b_{\ell-1})}} { 1+e^{\theta+b_{\ell-1}} \over 1+e^{\theta+b_{\ell}}},\
{ 1\over e^{(b_{\ell+1}-b_{\ell})}-1} { 1+e^{\theta+b_{\ell+1}} \over 1+e^{\theta+b_{\ell}}}
\right\}>0,
\\
D_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} {1\over (1+e^{\theta+b_\ell})(1+e^{-(\theta+b_\ell)})}>0.
\end{align}
\end{rmk}

\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$. Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}. Define $r_{\max}=\max_k r_k$. Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
\[
\text{MSE}(\hat \Theta, \trueT)\leq {c_1U_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
\]
and
\[
\text{MSE}(\hat\mb, \trueb) \leq  {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
\]
where $c_1,c_2>0$ are two constants independent of the tensor dimension and rank, and $U_{\alpha,\beta,\Delta}$, $L_{\alpha,\beta,\Delta}$, $C_{\alpha,\beta,\Delta}$, $D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
\end{thm}

\begin{rmk} It may seem counter-intuitive that the MSE of $\hat \Theta$ shows no dependence of $\hat \mb$ (and vice versa). In fact, from the proof of Theorem~\ref{thm:ratejoint}, the uniform constants $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}$ can be replaced by
\[
U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)\over g_\ell(\theta)|},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
\]
Therefore, the accuracy of $\hat \Theta$ depends on the accuracy of $\hat \mb$ implicitly through $U_{\alpha, \hat \mb}$ and $ L_{\alpha,\hat \mb}$.

\end{rmk}

\begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$, 
we have the following inequalities:
\begin{equation}\label{eq:level}
\flogl(\hat \Theta,\hat \mb)\geq \flogl(\trueT,\hat \mb)\quad \text{and}\quad \flogl(\hat \Theta,\hat \mb)\geq \flogl(\hat \Theta, \trueb).
\end{equation}

Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
\begin{equation}\label{eq:ratenew}
\text{MSE}(\hat \Theta, \trueT)\leq {CU_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
\end{equation}
where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two universal constants independent of the tensor dimension and rank.
Next we bound $\text{MSE}(\hat \mb,\trueb)$. It follows from the expression of $\flogl(\Theta,\mb)$ that
\begin{align}\label{eq:property}
{\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
\frac{\dot{f}(\theta_\omega+b_\ell)}{g_\ell(\theta_\omega)}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
\frac{\dot{f}(\theta_\omega+b_\ell)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
{\partial^2 \flogl\over \partial b_\ell^2}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(\theta_\omega+b_\ell)g_\ell(\theta_\omega)-\dot{f}(\theta_\omega+b_\ell)^2}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(\theta_\omega+b_\ell)g_{\ell+1}(\theta_\omega)+\dot{f}(\theta_\omega+b_\ell)^2}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
&\quad  \text{ for all } \ell \in [L-1],\notag \\
{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(\theta_\omega+b_\ell)\dot{f}(\theta_\omega+b_{\ell+1})}{g^2_{\ell+1}(\theta_\omega)}\quad \text{and}\quad 
\ {\partial^2 \flogl\over \partial b_\ell \partial b_\ell'} = 0 \text{ if } |\ell-\ell'|>1.
\end{align}

Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.

We consider the profile log-likelihood $\flogl(\mb, \hat \Theta)$ as a function of $\mb\in\tB$. For notational convenience, we drop $\hat \Theta$ from $\flogl(\mb, \hat \Theta)$ and simply write $\flogl(\mb)$. By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
\begin{equation}\label{eq:taylorb}
\flogl(\hat\mb)=\flogl(\trueb)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ {1\over 2}(\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
\end{equation}
where $\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.

The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
\begin{equation}\label{eq:linearb}
(\trueb-\hat\mb)^T\fploglb(\trueb)\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}
\leq \FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_\ell+n_{\ell+1}),
\end{equation}
where the last inequality is followed from
\[
\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{\mb=\trueb}\right| \leq C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_{\ell}+n_{\ell+1}), \quad \textrm{ for all } \; \ell\in [L-1].
\]
